---
title: "Tidyverse Reading and Manipulating Data"
format: html
editor: visual
---

```{r}
#|echo: false
library(dplyr)
library(readr)
library(purrr)
library(curl)
library(readxl)
```

## Task 1

1.  If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

> We could specify ../MyData.csv.

2.  What are the major benefits of using R projects?

> R projects ensure that your project working directory is correctly set and can keep tabs on your scripts and imports, essentially forming an encapsulated version of your work.

3.  What is git and what is github?

> Git is version control software that allows us to maintain a history of changes to code. Github is a service that maintains web-based repositories for us.

4.  What are the two main differences between a tibble and a data.frame?

> A tibble does "fancy" printing of data and does not coerce selection with \[\] into vectors.

5.  Rewrite the following nested function call using BaseR’s chaining operator:

```         
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species),
Petal.Length < 1.55), Species)
```

```{r}
as_tibble(iris) |> 
  select(starts_with("Petal"), Species) |> 
  filter(Petal.Length < 1.55) |> 
  arrange(Species)
```

## Task 2 Glass

read data and add column names

```{r}
glass_cols <- c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type of Glass")

glass_tbl <- read_csv("https://www4.stat.ncsu.edu/online/datasets/glass.data", col_names = glass_cols) 
```

create a factor to represent the types of glass, then pass to filter with our conditions

```{r}
glass_type_labels <- c(" building_windows_float_processed",
                       " building_windows_non_float_processed",
                       "vehicle_windows_float_processed",
                       "vehicle_windows_non_float_processed",
                       "containers",
                       "tableware",
                       "headlamps")

glass_tbl |> 
  mutate(`Type of Glass` = factor(`Type of Glass`, levels = seq(1:7), labels = glass_type_labels)) |> 
  filter(Fe < .2 & (`Type of Glass` == "tableware" | `Type of Glass` == "headlamps"))
```

## Task 2 Yeast

read, add column names, and remove undesired columns

```{r}
yeast_cols <- c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class")
yeast_col_widths <- c(12, 6, 6, 6, 6, 6, 6, 6, 6, 3)

yeast_dl <- curl("https://www4.stat.ncsu.edu/online/datasets/yeast.data")
yeast_tbl <- read_fwf("https://www4.stat.ncsu.edu/online/datasets/yeast.data",
                      col_positions = fwf_widths(widths = yeast_col_widths, col_names = yeast_cols)) |> 
  select(-seq_name, -nuc)
```

add mean and medians by class

```{r}
yeast_tbl |> 
  group_by(class) |> 
  mutate(across(where(is.numeric), list(mean = mean, med = median), .names = "{.col}_{.fn}"))
```

## Task 3

import and print white wine data

```{r}
white_wine_tbl <- read_xlsx("./data/white-wine.xlsx", sheet = 1)
white_wine_tbl
```

read names and apply to previous tibble, then add white column

```{r}
white_wine_cols <- read_xlsx("./data/white-wine.xlsx", sheet = 2) |> 
  pull(1)

white_wine_tbl <- white_wine_tbl |> 
  set_names(white_wine_cols) |> 
  mutate(color = "white")

white_wine_tbl
```

read and replace variable names as above, but for red wine this time

```{r}
red_wine_tbl <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", col_names = FALSE, delim = ";", skip = 1)

red_wine_tbl <- red_wine_tbl |> 
  set_names(white_wine_cols) |> 
  mutate(color = "red")

red_wine_tbl
```

now we combine and use a chain to filter, sort, select, and summarize

```{r}
wine_tbl <- bind_rows(red_wine_tbl, white_wine_tbl) |> 
  filter(quality > 6.5 & alcohol < 132) |> 
  arrange(desc(quality)) |> 
  select(contains("acid"), alcohol, color, quality) |> 
  group_by(quality) |> 
  mutate(qual_grp_alc_mean = mean(alcohol), qual_grp_alc_sd = sd(alcohol))

wine_tbl
```

